---
title: "detection_of_IoT_botnet_attacks_N_BaIoT Data Set exploration"
author: "Yulia Zamyatina"
date: "April, 2021"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE, message = FALSE, warning = FALSE}
# Install and load required libraries
if ( !require( tidyverse )) install.packages( "tidyverse", repos = "http://cran.us.r-project.org" )
library( tidyverse )
library( rvest )
library( ggplot2 )
library( tools )

# Folder with all IoT attacks data files from all devices
data_folder <- "data"

# Variable for data frame
df <- NULL

# Read data from the sampler data file, that contains 1000 rows for each 'botnet'
readDataSample <- function() {
  
  # All IoT attacks data gathered in one file
  data_file <- file.path( data_folder, "data_sample.csv" )
  
  # Data file url on GitHub
  datafile_url <- "https://github.com/juliazam/IoT-attack-detection/raw/master/data/data_sample.csv"
  
  # If data file doesn't exist in local folder, download it 
  if( ! file.exists( data_file ) ) {
    
    res <- download.file( datafile_url, data_file, mode = "wb" )
    
  }
  
  print( "Reading data from data file.")
  df <- read.csv( data_file, as.is = TRUE )
  
  return( df )
}

# Read existing data
df <- readDataSample()

colnames <- colnames( df )
```

# Data set description
  
According to UCI Machine Learning Repository [^1], this data set is the collection of real traffic data, gathered from 9 commercial IoT (*Internet of Things*) devices authentically infected by Mirai and BASHLITE (Gafgyt).

The data set has 115 attributes (parameters), below is the description of their headers:

1. It has 5 time-frames: L5 (1 min), L3 (10 sec), L1 (1.5 sec), L0.1 (500 ms) and L0.01 (100 ms)[^2].
2. The statistics extracted from each traffic stream for each time-frame:
     - *weight*: the weight of the stream (can be viewed as the number of items observed in recent history)
     - *mean*
     - *std (variance)*
     - *radius*: the root squared sum of the two streams' variances
     - *magnitude*: the root squared sum of the two streams' means 
     - *covariance*: an approximated covariance between two streams
     - *pcc*: an approximated correlation coefficient between two streams
3. It has following stream aggregations:
     - *MI*: ("Source MAC-IP" in N-BaIoT paper) Stats summarizing the recent traffic from this packet's host (IP + MAC)
     - *H*: ("Source IP" in N-BaIoT paper) Stats summarizing the recent traffic from this packet's host (IP)
     - *HH*: ("Channel" in N-BaIoT paper) Stats summarizing the recent traffic going from this packet's host (IP) 
           to the packet's destination host.
     - *HH_jit*: ("Channel jitter" in N-BaIoT paper) Stats summarizing the jitter of the traffic going from this packet's             host (IP) to the packet's destination host.
     - *HpHp*: ("Socket" in N-BaIoT paper) Stats summarizing the recent traffic going from this packet's host+port (IP) 
           to the packet's destination host+port. Example 192.168.4.2:1242 -> 192.168.4.12:80
  
Thus, the column *'MI_dir_L5_weight'* in the data set shows the weight of the recent traffic from the packet's host for L5 time-frame.
  
The data set consists of \*.csv files, each representing a benign traffic or an attack. When I gathered \*.csv files together in one data set, I added *'botnet'* column, where I keep information about the attacks from the different botnets. The dataset contains *combo*, *junk*, *scan*, *tcp* and *udp* Gafgyt attacks, and *ack*, *scan*, *syn*, *udp* and *udpplain* Mirai attacks. I used *'ga'* prefix for Gafgyt attacks and *'ma'* for Mirai attacks in the *'botnet'* column.

List of attacks can be found in "N-BaIoT: Network-based Detectionof IoT Botnet Attacks Using Deep Autoencoders"[^3] article.

# Data set exploration

Below I explore the data only from *Danmini doorbell* device. Because the whole dataset is huge, I'll use a sample data set that contains about 1000 rows for each botnet just for illustrative purpose. Otherwise, plots will be too dense.

## MI stream

The few first columns contain the data for *MI* stream, and I start my research from *weight* data for L5 - L0.01 time-frames (MI_dir_L5_weight - MI_dir_L0.01_weight columns):

```{r echo = FALSE, message = FALSE, warning = FALSE, width = 8}
pattern <- "MI_dir_L(0)*(.)?(0)*\\d_(we)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
The plot shows that using only the *weight* attribute, I can easily separate **benign traffic**, **ga_tcp** and **ga_udp** attacks from the other attacks. Boxplots for **benign traffic**, **ga_tcp** and **ga_udp** shows that their medians are close to 0, they do not have large IQR, they do not have outliers. 

This is more clearly seen at the small time-frames, let's see the close up of the *weight* attribute for L1 and L0.01 time-frames:

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
df %>% group_by( botnet ) %>% select (MI_dir_L1_weight ) %>% 
  ggplot( aes( MI_dir_L1_weight, botnet,  fill = botnet ) ) + 
  geom_boxplot() +
  theme_minimal()
  
df %>% group_by( botnet ) %>% select (MI_dir_L0.01_weight ) %>% 
  ggplot( aes( MI_dir_L0.01_weight, botnet,  fill = botnet ) ) + 
  geom_boxplot() +
  theme_minimal()
```

The **ga_tcp** and **ga_udp** disguise themselves well as **benign traffic**. Thus, I have to find at least one other attribute, that can help me separate **benign traffic** from **ga_tcp** and **ga_udp** attacks.

Next, I would like to explore *mean* attribute for the same stream (MI_dir_L5_mean -  MI_dir_L0.01_mean):

```{r echo = FALSE, message = FALSE, warning = FALSE}
pattern <- "MI_dir_L(0)*(.)?(0)*\\d_(me)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
This plot shows that *mean* attribute can help me separate **benign traffic** from **ga_tcp** and **ga_udp**, as median for **benign traffic** is higher that for these attacks. Let's check this on the small time-frames (L1 and L0.01):
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
df %>% group_by( botnet ) %>% select (MI_dir_L1_mean ) %>% 
  ggplot( aes( MI_dir_L1_mean, botnet,  fill = botnet ) ) + 
  geom_boxplot() +
  theme_minimal()
  
df %>% group_by( botnet ) %>% select (MI_dir_L0.01_mean ) %>% 
  ggplot( aes( MI_dir_L0.01_mean, botnet,  fill = botnet ) ) +
  geom_boxplot() +
  theme_minimal()
```
  
Some outliers for *benign traffic* have values close to those of **ga_tcp** and **ga_udp**, and these values may interfere with separation of *benign traffic* from attacks. Thus, using only the *weight* and *mean* attributes will not help to clearly separate **benign traffic** from the attacks. Therefore, I still need one more attribute that helps separate *benign traffic* from the **ga_tcp** and **ga_udp** attacks. 

Next, let's explore *variance* attribute (MI_dir_L5_variance - MI_dir_L0.01_variance) - does it provide any information for separating *benign traffic* from the attacks:

```{r echo = FALSE, message = FALSE, warning = FALSE}
pattern <- "MI_dir_L(0)*(.)?(0)*\\d_(var)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
This plot shows that *variance* attribute doesn't give new information on how to separate **benign traffic** from attacks, so I can easily remove this attribute if I need to reduce the data set dimensions.

## H and HH streams
  
Let's explore statistics for H and HH streams. Similarly to the previous stream, I will consider *weight*, *mean* and *variance* (or *std*) attributes:
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 8}
# Weight
pattern <- "H_L(0)*(.)?(0)*\\d_(we)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
  
# Mean
pattern <- "H_L(0)*(.)?(0)*\\d_(me)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
# Variance
pattern <- "H_L(0)*(.)?(0)*\\d_(var)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
    facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
    theme( axis.text.x = element_blank(), legend.position = "bottom" )
  
# Std
pattern <- "H_L(0)*(.)?(0)*\\d_(std)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
Just like plots for MI stream, all plots for H and HH streams show that I can use *weight* and *mean* attributes to separate **benign traffic** from the attacks, and remove *variance* or *std* attributes if I need to reduce the data set dimensions.

### HH stream

HH stream also has *magnitude*, *radius*, *covariace* and *pcc* attributes, let's explore them:
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
pattern <- "H_L(0)*(.)?(0)*\\d_(ma)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
The plot shows that *magnitude* attribute can be used to separate **benign traffic** from the attacks.
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
pattern <- "H_L(0)*(.)?(0)*\\d_(rad)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
The plot shows that *radius* attribute doesn't give any information on how to separate **benign traffic** from the attacks and can be removed if I need to reduce data set dimensions.

```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
pattern <- "H_L(0)*(.)?(0)*\\d_(co)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
*Covariance* attribute can be used to separate attacks from **benign traffic**, and likely it's the one I've looked for.
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
pattern <- "H_L(0)*(.)?(0)*\\d_(pcc)"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
The plot shows that *pcc* attribute can also be used for separation.

## HH_jit stream

I will follow the same pattern in exploring attributes for this stream as for the previous ones:
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 7}
pattern <- "HH_jit"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 3 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```
  
These plots have no new information on how to separate **benign traffic** from attacks.

## HpHp stream
  
Let's have a quick look at plots for this stream:
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 15, fig.width = 12}
pattern <- "HpHp"
  
tmp <- df %>% select( botnet, grep( pattern = pattern, ignore.case = TRUE, x = colnames ) )
  
tmp %>% gather( attacks, data, -botnet ) %>% 
  ggplot( aes( botnet, data, fill = botnet ) ) +
  geom_boxplot() +
  facet_wrap( ~attacks, scales = "free", ncol = 5 ) +
  theme( axis.text.x = element_blank(), legend.position = "bottom" )
```

This stream also doesn't give me any new information on how to separate **benign traffic** from attacks.

## Data set exploration at small time frames
  
I've noticed that all the plots have more pronounced data at the small time-frames. I would like to explore which combination of attributes, for example *weight* vs. *mean*, will give me information on how to separate **benign traffic** from attacks on these small time-frames.
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
df %>% ggplot( aes( MI_dir_L1_weight, MI_dir_L1_mean, color = botnet ) ) + geom_point()
df %>% ggplot( aes( MI_dir_L0.01_weight, MI_dir_L0.01_mean, color = botnet ) ) + geom_point()
```
  
Pair *weight-mean* for L0.01 time-frame shows how easily some attacks can be separated from **benign traffic**, compared with L1 time-frame.
  
```{r echo = FALSE, message = FALSE, warning = FALSE, fig.height = 4}
df %>% ggplot( aes( HH_L1_mean, HH_L1_covariance, color = botnet ) ) + geom_point()
df %>% ggplot( aes( HH_L0.01_mean, HH_L0.01_covariance, color = botnet ) ) + geom_point()

rm( df, tmp, pattern, error, colnames, data_file, data_folder, data_url, sample_size )
```

*mean-covariance* pair better shows how **ga_tcp** and **ga_udp** can be separated from **benign traffic**.

# Conclusion

As a result, all explorations show that I can use *weight*, *mean* and *covariance* attributes to make a decision on how to separate **benign traffic** from attacks, and remove other statistic attributes if I need to reduce the data set dimensions.

[^1]: https://archive.ics.uci.edu/ml/datasets/detection_of_IoT_botnet_attacks_N_BaIoT
[^2]: https://arxiv.org/pdf/1805.03409.pdf, p.3
[^3]: https://arxiv.org/pdf/1805.03409.pdf, p.5